\documentclass[%
 sor,
%aip,
%twoside,
%groupedaddress,
%jmp,
 jor,
 amsmath,amssymb,
%preprint,%
 reprint,%
%author-year,%
%author-numerical,%
]{revtex4-2}

\usepackage{graphicx}% Include figure files
\usepackage{xcolor}
\usepackage{circuitikz}
\usepackage{siunitx}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{amsmath, amssymb, amsfonts}
\usepackage{placeins}
\usepackage{float}
\usepackage{gensymb}
\usepackage{array}
\usepackage{siunitx}
\usepackage{enumitem}

\begin{document}

\preprint{IDK what this does}

\title{Experiment 4\\Specific Heat of Solids}

\author{Aumshree P. Shah\\20231059\color{red}}
\altaffiliation[\color{red}]{aumshree.pinkalbenshah@students.iiserpune.ac.in}
\date{5 Feb 2025}
\vspace{1cm}
\begin{abstract}
\centering
In this experiment we try to measure specific heat capacity of solids.
\end{abstract}
\maketitle
\section{Theory and Procedure}
\subsection{Apparatus}

{\small
\begin{itemize}
\begin{minipage}[t]{0.45\textwidth}
    \item Calorimeter
    \item Shots of material(s)
    \item Weighing balance
    \item Steam chamber
\end{minipage}
\hfill
\begin{minipage}[t]{0.45\textwidth}
    \item Steam generator
    \item Electric heater
    \item Thermometer
\end{minipage}
\end{itemize}
}

\subsection{Theory}

The change in temperature $\Delta T$ of a material when supplied with a fixed amount of heat $\Delta Q$ depends
on the type of material and is inversely proportional to the the mass $m$ of the material. The material dependence
is given the quantity called the specific heat $c$ of the substance. The interdependence of these quantities is
summarised by the equation$$\Delta Q= cm \Delta T$$

The amount of heat required to raise the temperature of a unit mass of a substance by 1\si{\celsius} is called the
specific heat of the substance. The SI unit of specific heat is [\si{\joule\per\kilogram\per\kelvin}]. The specific heat of water is among the highest of all substances. Historically, the specific heat of water is arbitrarily set at 1cal \si{\per\gram\per\celsius}. The equivalent SI value is 4184 \si{\joule\per\kilogram\per\kelvin}. A related quantity is the molar specific heat $c_m$, which is the specific heat for 1  \si{\mole} of a substance. \\

If $c_s$ is the heat capacity of shots, $\Delta T_s \equiv T_f -T_s$ is the change in temperature of shorts and $m_s$ is the mass of shorts, then the heat lost by the shots is
$$\Delta Q_s = c_s m_s \Delta T_s$$
Similarly, heat gained by the water is
$$\Delta Q_w = c_w m_ w\Delta T_w$$
with $c_w$, $m_w$, $T_w$ and $\Delta T_w$ have their proper meanings.  \\ 

Now, if there is no heat absorbed by the calorimeter, then $\Delta Q_w = \Delta Q_s$, but that is not the case. To make the equation equal, we replace $m_w$ by $m_w+m_f$ where $m_f$ is the equivalent water mass of the calorimeter. Hence we get
\begin{equation}
c_s = \frac{c_w(m_w +m_f)\Delta T_w}{m_s \Delta T_s}
\end{equation} \\

\subsection{Procedure}
First we measure the mass of the shots of each material. Then we place shots of material into the steam chamber and close them using a cork with hole through which we put a thermometer to measure the $T_s$. We raise the temperature of steam generator using electric heater. When the temperature reaches equilibrium, we turn the chamber over and quickly pour all the material into a calorimeter with measured mass and temperature of water. We stir it using a different thermometer and measure the equilibrium temperature $T_f$. \\
We repeat this same procedure for different materials and then to measure $m_f$ of the calorimeter, we put water of higher temperature $\theta_{i1} \approx 60$\si{\celsius}\footnote{this is to reduce loss in evaporation and other convective losses} into room temperature water inside calorimeter and measuring the equilibrium temperature $\theta_f$; using the same principle:
\begin{equation}
m_f = \frac{m_1\Delta \theta_1 - m_2 \Delta \theta_2}{\Delta \theta_2}
\end{equation} \\

\section{Observations}



\begin{table}[ht]
\centering
    \begin{minipage}[b]{0.48\hsize}\centering
\begin{tabular}{|c|c|c|c|c|}
\hline
\textbf{Material} & \textbf{$m_s$ (\si{\gram})} & \textbf{$m_w$ (\si{\gram})} & \textbf{$T_w$ (\si{\celsius})} & \textbf{$T_f$ (\si{\celsius})} \\
\hline
Copper & 100.1 & 90.0 & 23.5 & 30.1 \\
Acrylic & 44.5 & 45.0 & 24.9 & 35.0 \\ % Fixed spelling
Copper\footnote{Took the copper out earlier than equilibrium temperature} & 183.5 & 52.0 & 23.0 & 39.1 \\ 
Ceramic & 31.4 & 69.0 & 23.9 & 29.3 \\
\hline
\end{tabular}
\caption{With calorimeter of type-1, data was taken on 23 Jan 2025}
\label{tab:table1}
\vspace{0.5cm}Data for Type 1:
\[
\boxed{
\begin{aligned}
m_2 &= 62.2~\si{\gram} \\
T_2 &= 20.7~\si{\celsius} \\
m_1 &= 41.5~\si{\gram} \\
T_1 &= 56.6~\si{\celsius} \\
T_f &= 33.8~\si{\celsius}
\end{aligned}
}
\]

    \end{minipage}
\hfill\vline\hfill
    \begin{minipage}[b]{0.48\hsize}\centering
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\textbf{Material} & \textbf{$m_s$ (\si{\gram})} & \textbf{$m_w$ (\si{\gram})} & \textbf{$T_s$ (\si{\celsius})} & \textbf{$T_w$ (\si{\celsius})} & \textbf{$T_f$ (\si{\celsius})} \\
\hline
Copper & 55.0 & 176.4 & \multicolumn{1}{c}{--} & 26.1 & 30.0 \\ % Centered dash
Copper\footnote{Spillage while pouring water} & 171.1 & 154.0 & 90--91 & 25.1 & 31.2 \\
Acrylic & 52.3 & 144.4 & 87--88 & 25.1 & 29.3 \\ % Fixed spelling
Ceramic & 39.6 & 132.4 & 80 & 24.8 & 28.6 \\
Copper & 180.0 & 180.0 & \multicolumn{1}{c}{--} & 22.3 & 28.5 \\
\hline
\end{tabular}
\caption{With calorimeter of type-2, data taken on 29 Jan 2025}
\label{tab:table2}
\vspace{0.5cm}Data for Type 2:\\
\[
\boxed{
\begin{aligned}
m_2 &= 129.7~\si{\gram} \\
T_2 &= 19.2~\si{\celsius} \\
m_1 &= 87.5~\si{\gram} \\
T_1 &= 56.6~\si{\celsius} \\
T_f &= 33.8~\si{\celsius}
\end{aligned}
}
\]
    \end{minipage}
\end{table}





\section{Uncertainties and Sources of Error}
\subsection{Uncertainties from precision of instrument}
\begin{itemize}
    \item Weight Measurements:
    All weight values have an uncertainty of $\Delta m = \pm 0.05$ \si{\gram} due to instrument resolution.    
    \item Temperature Measurements: Temperature values have an uncertainty of $\Delta T = \pm 0.005$ \si{\kelvin} due to instrument resolution. 
\item Mercury Thermometer: The thermometer used to measure $T_s$ has an uncertainty of $\Delta T_m = \pm 0.5$ \si{\kelvin}
\end{itemize}

\subsection{Systematic Errors and Precautions}  
Potential systematic errors included spillage during transfers, residual material retention in containers, unintended material escape from the calorimeter, heat dissipation through radiation (according to Newton's law of cooling), assumptions of instrument calibration accuracy, post-experiment temperature inconsistencies in the calorimeter due to inadequate calibration, and uneven heating.  

Precautionary measures included preheating reused materials to eliminate residual moisture, repeating tests with procedural anomalies to replace outliers, and maintaining calorimeter thermal stability by refilling them with tap water after each run to standardize initial conditions.


\section{Calculation and Error Analysis}
From Table-\ref{tab:table1} and Table-{\ref{tab:table2}} we discard the two copper data points which had some error when performing before doing the calculation. 
\subsection{Error Propagation}

The heat capacity of copper is given by Equation-1. The uncertanity in $c_s$ is given by the equation:
$$
   \frac {\Delta c_s} {c_s} =\sqrt{ 
(\partial m_w \Delta m)^2 + (\partial m_f \Delta m)^2 + (\partial m_s \Delta m)^2 + (\partial T_w \Delta T)^2 + (\partial T_f \Delta T)^2 + (\partial T_s \Delta T_m)^2 
}$$
Where these partials are calculated by\\
\begin{minipage}{0.5\textwidth}
\begin{align*}
  {\partial m_w} &= \frac{c_w \Delta T_w}{m_s \Delta T_s} \\
  {\partial m_s} &= \frac{c_w(m_w +m_f)\Delta T_w}{m_s^2 (t_s - t_f)} \\
    {\partial m_f} &= \frac{c_w \Delta T_w}{m_s \Delta T_s}
\end{align*}
\end{minipage}%
\begin{minipage}{0.5\textwidth}
\begin{align*}
{\partial T_w} &= \frac{(m_w + m_s) c_w}{m_s} {\Delta T_s} \\
   {\partial T_f} &= \frac{(m_w + m_f) c_w \Delta T_w)}{m_s \Delta T_w^2} \\
    {\partial T_s} &= \frac{c_w(m_w +m_f)\Delta T_w}{m_s \Delta T_s^2}
\end{align*}
\end{minipage}
\subsection{Calculation}
For non observed data potms we take sam einital temperatuer. disdcaard twwo non coper data points and take +- 3 grams of errpr of water equivalent for cimplicity. 

\section{Result}
\ref{github}

The reference resistance was determined to be: $R_0 = 2.90 \pm 7.7\% $. (This error accounts for uncertainty in temperature measurements)\\

The Stefan-Boltzmann fit for the tungsten filament showed:
$$
P = (4.00 \pm 0.066) \times 10^{-14}\ \mathrm{T^4}
$$

The log-log analysis revealed a power-law relationship:
$$
\log P = (3.98 \pm 0.06)\log T + (-30.71 \pm 0.48)
$$

\noindent\fbox{%
    \parbox{\textwidth}{%
       The measured slope of $m = 3.98 \pm 0.06$ agrees with Stefan's law prediction of $P \propto T^4$, as the theoretical value of 4 lies within the experimental uncertainty range.
       }%
}\\
\\

\noindent\rule{\linewidth}{0.4pt}
\vspace{2cm}
\appendix
\section{water equivilane}\label{appendix:prevexp}
ONly single measurement of wateer equivilant taken.
\section{Ambient Temperature}
No measure of ambient temperature is taken in the latter experiment, since it had no use in calculations other than the fact that it can be neglected. 

\bibliography{aipsamp}
\end{document}





